Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
update results section

For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors. Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$. Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.

@@

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together. The highest value for each performance metric
is highlighted in bold.}}
\label{tab:result_weighted_avg}
 \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
 MLP & \textbf{0.814} & 0.732 & 0.771 & 0.987 & 0.987 & 0.987 \\ \hline
 BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
 CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.

Table~\ref{tab:result_roc} shows the impact of individual fetures as well as their combination on the segmentation task. Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.\\

\begin{table}[ht]
\centering
\caption{\textbf{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used. Highest value for each experiment across all models is highlighted in boldface.}}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
 & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 word embeddings + POS & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 word embeddings + punctuation & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}

>>>>

For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors. Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$. From the results illustrated in Figure~\ref{fig:length-mlp}, we observed that best performance is achieved when $l = 2$. Therefore, all MLP results are reported with $l = 2$ in the remaining part of this section.

@@

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together. The highest value for each performance metric
is highlighted in bold.}}
\label{tab:result_weighted_avg}
 \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
 MLP & \textbf{0.817} & 0.710 & 0.760 & 0.986 & 0.987 & 0.986 \\ \hline
 BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
 CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together. Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.

Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task. Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively. Individually, POS and punctuation features improve the performance of all classifiers except CRNN when word embeddings are used with POS features. CRF and MLP achieve their highest AUCs when all features are used together. On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features. \\

\begin{table}[ht]
\centering
\caption{\textbf{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used. Highest value for each experiment across all models is highlighted in boldface.}}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
 & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 word embeddings + POS & \textbf{0.973} & 0.953 & 0.920 & 0.962 \\ \hline
 word embeddings + punctuation & \textbf{0.993} & 0.989 & 0.982 & 0.988 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}
